### 2018-04-04 Maths Notes/Statistics/S3/Estimation, confidence intervals and tests/Estimation.tex
S3 Notes Update

\begin{document}
\underline{\huge Estimation, confidence intervals and tests}

@@

If $E(T)=\theta$ then the statistic is unbiased

>>>>

\begin{document}
\begin{center}
\underline{\huge Estimation, confidence intervals and tests}
\end{center}
\section{Statistics}
A statistic is defined as follows:\\
If $X_1,X_2,X_3,...,X_n$ is a random sample of size n from some population, then a statistic T is a random variable consisting of the $X_i$ than involves no other quantities.\\
\\
Since it is possible to repeat the process of taking a sample, if all samples are taken, they will form a probability distribution called the \textbf{sampling distribution of T}. This will usually depend on the distribution of the population X.\\
\\
A random sample is defined as follows:\\
A random sample of size n consists of the observations $X_1,X_2,X_3,...,X_n$ from a population where the $X_i$
\begin{itemize}
\item Are independent random variables
\item Have the same distribution as the population
\end{itemize}

@@

If $E(T)=\theta$ then the statistic is unbiased\\
{\large $\mathlarger{\hat{\theta}}$} represents an estimator of $\theta$, this is true for all population parameters, such as $\hat{\mu}$

### 2018-04-04 Maths Notes/Statistics/S3/Exam Questions/Confidence Intervals/Confidence Intervals.tex
S3 Notes Update

\begin{document}
\begin{center}
\underline{\huge Exam Questions - Combinations of random variables}
\end{center}
\section{Example 1}
\textit{Philip and James are racing car drivers. Philip’s lap times, in seconds, are normally distributed with mean 90 and variance 9. James’ lap times, in seconds, are normally distributed with mean 91 and variance 12. The lap times of Philip and James are independent. Before a race, they each take a qualifying lap.\\
\\
Find the probability that James’ time for the qualifying lap is less than Philip’s.}\\
\textcolor{red}{Write down the two distributions}
$$P\sim N(90,9) \qquad J\sim N(91,12)$$
\textcolor{red}{Write the probability wanted to be found, and rearrange so the distributions are on the same side}
$$P(J<P)$$
$$P(J-P<0)$$
\textcolor{red}{Combine the two distributions}
$$(J-P)\sim N(1,21)$$
\textcolor{red}{Find the z value given the new distribution}
$$P(Z<\frac{0-1}{\sqrt{21}}$$
$$P(Z<-0.2182)$$
\textcolor{red}{Convert into a probability using the tables}
$$1-\Phi(0.2182)$$
$$1-0.5871=0.4129$$
\newpage
\section{Example 2}
\textit{The random variable A is defined as}
$$A=4X-3Y$$
\textit{where $X\sim N(30,3^2)$, $Y\sim N(20,2^2)$ and X and Y are independent\\
The random variables $Y_1$, $Y_2$, $Y_3$ and $Y_4$ are independent and each has the same distribution as Y.\\
The random variable B is defined as}
$$B=\sum^4_{i=1}Y_i$$
\textit{Find $P(B>A)$}\\
\textcolor{red}{Find the distributions of A and B}
$$E(A)=E(4X-3Y)=4E(X)-3E(Y)=4\times30-3\times20=60$$
$$Var(A)=Var(4X-3Y)=16Var(X)+9Var(Y)=16\times9+9\times4=180$$
$$E(B)=4E(Y)=4\times20=80$$
$$Var(B)=4Var(Y)=4\times4=16$$
\textcolor{red}{Rewrite the probability to get both distributions on the same side}
$$P(B-A>0)$$
\textcolor{red}{Combine A and B to form the distribution B-A}
$$E(B-A)=80-60=20$$
$$Var(B-A)=180+16=196$$
\textcolor{red}{Find the z value given the new distribution}
$$P(Z>\frac{-20}{\sqrt{196}}$$
$$P(Z>-1.428)$$
$$\Phi(1.428)=0.9236$$

>>>>

\begin{document}
\begin{center}
\underline{\huge Exam Questions - Confidence Intervals}
\end{center}
\section{Normal mean}
\textit{A woodwork teacher measures the width, w mm, of a board. The measured width, x mm, is
normally distributed with mean w mm and standard deviation 0.5 mm.\\
The same board is measured 16 times and the results are recorded.\\
Given that the mean of these 16 measurements is 35.6 mm,\\
find a 98\% confidence interval for w.}
\textcolor{red}{Find the standard error}
$$\textrm{Standard error}=\frac{\sigma}{\sqrt{n}}=\frac{0.5}{\sqrt{16}}=\frac{1}{8}$$
\textcolor{red}{Find the z greater value for 1\%}
$$P(Z>2.3263)=0.01$$
\textcolor{red}{Substitute in the values}
$$\overline{x}\pm z\times\textrm{Standard error}$$
$$35.6\pm2.3263\times\frac{1}{8}$$
